---
title: "CEE 272R Case 2 Exploration"
author: "Sam Kramer"
date: "6/3/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(forcats)
library(stringr)

path_fig = "~/Desktop/"
```

## Functions

This function rounds battery sizes to the nearest 0.005, or 50 MWh. 
```{r}
round_frac <- function(x) {
  x <- x * 200
  x <- ceiling(x)
  x <- x / 200
  x
}
```


## Read data
```{r}
setwd("~/Documents/School/Spring/CEE 272R/Project/CEE272R/case_2/code/")

up_ramps_raw <- read_csv("../results/battery_power_up.csv",
                         col_names = FALSE)

down_ramps_raw <- read_csv("../results/battery_power_down.csv",
                           col_names = FALSE)

ramp_rates_up_raw <- read_csv("../results/Ramp_Rates_Up.csv",
                              col_names = FALSE)
ramp_rates_down_raw <- read_csv("../results/Ramp_Rates_Down.csv",
                                col_names = FALSE)

SCALE_FACTOR = 10000
```

## Clean Data

### Clean battery sizing data
```{r}
# Create column names (i.e. ramp reduction amounts) for the data frame
reduction_up <- seq(0,(ncol(up_ramps_raw) - 1) * 0.0001,
                    length.out = ncol(up_ramps_raw))
colnames(up_ramps_raw) <- reduction_up

reduction_down <- seq(0,(ncol(down_ramps_raw) - 1) * 0.0001,
                      length.out = ncol(down_ramps_raw))
colnames(down_ramps_raw) <- reduction_down

# Add a column for "event number" and tidy the data
up_ramps <- up_ramps_raw %>% 
  mutate(event_number = row_number()) %>%
  # mutate(event_number = row_number()) %>% 
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("u", event_number)) %>% 
  select(-event_number)
         # direction = "up")

up_ramps

down_ramps <- down_ramps_raw %>% 
  mutate(event_number = row_number()) %>%
  # mutate(event_number = row_number()) %>% 
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("d", event_number)) %>% 
  select(-event_number)
         # direction = "down")

down_ramps
```

### Clean true ramp rate data

#### Up Ramps
```{r}
ramp_rates_up <- ramp_rates_up_raw

colnames(ramp_rates_up) <- c("ramp_rate")

ramp_rates_up <- ramp_rates_up %>%
  mutate(event_number = row_number(),
         id = str_c("u", event_number)) %>% 
  # mutate(event_number = row_number())
         # direction = "up") %>% 
  # select(direction, event_number, ramp_rate, everything())
  select(-event_number)

ramp_rates_up
```

#### Down Ramps
```{r}
ramp_rates_down <- ramp_rates_down_raw

colnames(ramp_rates_down) <- c("ramp_rate")

ramp_rates_down <- ramp_rates_down %>%
  mutate(event_number = row_number(),
         id = str_c("d", event_number)) %>% 
  # mutate(event_number = row_number())
         # direction = "down") %>% 
  # select(event_number, ramp_rate)
  select(-event_number)

ramp_rates_down
```

## Create Table for All Ramps

```{r}
all_ramps <- up_ramps %>% 
  rename(up = battery_size) %>% 
  full_join(down_ramps %>% rename(down = battery_size),
            by = c("id", "slope_reduction")) %>% 
  gather(up, down, key = "direction", value = "battery_size")

all_ramps
```


```{r}
all_ramps <- up_ramps %>%
  rename(up = battery_size) %>%
  full_join(down_ramps %>%
              rename(down = battery_size),
            by = c("event_number","slope_reduction")) %>%
  gather(up, down, key = "direction", value = "battery_size") %>%
  select(direction, event_number, slope_reduction, battery_size)

all_ramps
```

```{r}
ramp_rates_all <- ramp_rates_up %>% 
  rename(up = ramp_rate) %>% 
  full_join(ramp_rates_down %>% rename(down = ramp_rate),
            by = "event_number") %>% 
  gather(-event_number, key = "direction", value = "ramp_rate")

ramp_rates_all
```


## Add variables

### Calculate percentage reduction

#### Up Ramps
```{r}
up_ramps <- up_ramps %>%
  left_join(ramp_rates_up, by = "event_number") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate) %>%
  select(event_number, battery_size, slope_reduction, percent_reduction)

up_ramps
```

#### Down Ramps
```{r}
down_ramps <- down_ramps %>% 
  left_join(ramp_rates_down, by = "event_number") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate) %>%
  select(event_number, battery_size, slope_reduction, percent_reduction)

down_ramps
```


#### All Ramps
```{r}
all_ramps <- all_ramps %>% 
  left_join(ramp_rates_all, 
            by = c("event_number", "direction")) %>% 
  filter(!is.na(ramp_rate)) %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate) %>% 
  select(event_number, direction, battery_size, slope_reduction, percent_reduction)

all_ramps
```


### Round battery sizes

#### Up Ramps
```{r}
# Round the battery sizes
up_ramps <- up_ramps %>% 
  mutate(rounded_size = round_frac(battery_size))

up_ramps
```

#### Down Ramps
```{r}
# Round the battery sizes
down_ramps <- down_ramps %>% 
  mutate(rounded_size = round_frac(battery_size))

down_ramps
```

#### All Ramps
```{r}
all_ramps <- all_ramps %>% 
  mutate(rounded_size = round_frac(battery_size))

all_ramps
```


### Find maximum slope reduction for each combination of event and battery size
```{r}
max_events_up <- up_ramps %>% 
  group_by(rounded_size,event_number) %>% 
  summarize(slope_reduction = max(slope_reduction),
            percent_reduction = max(percent_reduction)
            ) %>% 
  ungroup()

# max_events_percent <- up_ramps_percent %>% 
#   group_by(rounded_size,event_number) %>% 
#   summarize(percent_reduction = max(percent_reduction)) %>% 
#   ungroup()

max_events_up
# max_events_percent
```

```{r}
max_events_down <- down_ramps %>% 
  group_by(rounded_size,event_number) %>% 
  summarize(slope_reduction = max(slope_reduction),
            percent_reduction = max(percent_reduction)
            ) %>% 
  ungroup()

max_events_down
```


```{r}
max_events_all <- all_ramps %>% 
  group_by(rounded_size, event_number, direction) %>% 
  summarize(slope_reduction = max(slope_reduction),
            percent_reduction = max(percent_reduction)
            ) %>% 
  ungroup()

max_events_all
```

## Make full dataset for all battery sizes and events

#### Up Ramps
```{r}
full_dataset_up <- max_events_up %>% 
  data_grid(rounded_size, event_number) %>% 
  left_join(max_events_up, by = c("event_number", "rounded_size")) %>% 
  select(event_number, rounded_size, slope_reduction) %>% 
  arrange(event_number, rounded_size) %>% 
  mutate(slope_reduction = ifelse(rounded_size == 0, 0, slope_reduction)) %>% 
  fill(slope_reduction) %>% 
  left_join(ramp_rates_up, by = "event_number") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate)

full_dataset_up


full_medians_up <- full_dataset_up %>% 
  group_by(rounded_size) %>% 
  summarize(median = median(slope_reduction),
            median_percent = median(percent_reduction),
            q10 = quantile(slope_reduction, 0.1),
            q90 = quantile(slope_reduction, 0.9),
            q10_percent = quantile(percent_reduction, 0.1),
            q90_percent = quantile(percent_reduction, 0.9)
  )

full_medians_up
```

```{r}
full_dataset_down <- max_events_down %>% 
  data_grid(rounded_size, event_number) %>% 
  left_join(max_events_down, by = c("event_number", "rounded_size")) %>% 
  select(event_number, rounded_size, slope_reduction) %>% 
  arrange(event_number, rounded_size) %>% 
  mutate(slope_reduction = ifelse(rounded_size == 0, 0, slope_reduction)) %>% 
  fill(slope_reduction) %>% 
  left_join(ramp_rates_down, by = "event_number") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate)

full_dataset_down


full_medians_down <- full_dataset_down %>% 
  group_by(rounded_size) %>% 
  summarize(median = median(slope_reduction),
            median_percent = median(percent_reduction),
            q10 = quantile(slope_reduction, 0.1),
            q90 = quantile(slope_reduction, 0.9),
            q10_percent = quantile(percent_reduction, 0.1),
            q90_percent = quantile(percent_reduction, 0.9)
  )

full_medians_down
```

```{r}

```





## Data Exploration


## Plot data
```{r}
up_ramps %>% 
  ggplot(aes(battery_size * SCALE_FACTOR)) + 
  geom_histogram(binwidth = 0.05 * SCALE_FACTOR)

# up_ramps %>% 
#   filter(slope_reduction == 1*10^-4,
#          battery_size < quantile(battery_size, 0.9, na.rm = TRUE)) %>% 
#   ggplot(aes(battery_size)) + 
#   geom_histogram(bins = 100)

# up_ramps %>% 
#   filter(slope_reduction %in% quantile(up_ramps$slope_reduction)) %>% 
#   ggplot(aes(battery_size)) + 
#   geom_freqpoly(aes(color = factor(slope_reduction)), binwidth = 0.05)
```


```{r}
up_ramps %>% 
  filter(!is.nan(battery_size)) %>% 
  ggplot(aes(slope_reduction * SCALE_FACTOR)) + 
  geom_histogram(binwidth = 0.0001*SCALE_FACTOR) + 
  labs(x = "Ramp Rate Reduction (MW/hr)", y = "Count")

up_ramps %>% 
  filter(!is.nan(battery_size)) %>% 
  ggplot(aes(slope_reduction * SCALE_FACTOR, ..density..)) + 
  geom_histogram(binwidth = 0.0001*SCALE_FACTOR) + 
  labs(x = "Ramp Rate Reduction (MW/hr)", y = "Fraction of Events")

# up_ramps %>% 
#   filter(!is.nan(battery_size))
```


```{r}
full_dataset_up %>%
  filter(rounded_size <= 1.025) %>%
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  ggplot(aes(rounded_size, slope_reduction * SCALE_FACTOR)) +
  # geom_boxplot(aes(group = cut_number(rounded_size, 10)), outlier.alpha = 0) +
  geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0)), outlier.alpha = 0) +
  scale_x_continuous(breaks = c(0:10)) +
  coord_cartesian(xlim = c(1,10.5), ylim = c(0, 450)) + 
  labs(x = "Battery Size (GWh)", y = "Ramp Rate Reduction (MW/hr)") +
  theme_light()

full_dataset_up %>%
  filter(rounded_size <= 1.025) %>%
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  ggplot(aes(rounded_size, percent_reduction * 100)) +
  # geom_boxplot(aes(group = cut_number(rounded_size, 10))) +
  geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0)), outlier.alpha = 0) +
  coord_cartesian(xlim = c(1,10.5)) +
  scale_x_continuous(breaks = c(0:10)) +
  labs(x = "Battery Size (GWh)", y = "Ramp Rate Reduction (%)") +
  theme_light()

full_medians_up %>% 
  ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median * SCALE_FACTOR)) + 
  geom_ribbon(aes(ymin = q10 * SCALE_FACTOR, ymax = q90 * SCALE_FACTOR),
              alpha = 0.3) +
  geom_line() + 
  coord_cartesian(xlim = c(0, 50)) +
  labs(x = "Battery Size (GWh)", y = "Median Ramp Rate Reduction (MW/hr)") +
  theme_light()

full_medians_up %>% 
  ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median_percent * 100)) + 
  geom_ribbon(aes(ymin = q10_percent * 100, 
                  ymax = q90_percent * 100),
              alpha = 0.3) +
  geom_line() + 
  coord_cartesian(xlim = c(0,50)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Battery Size (GWh)", y = "Median Ramp Rate Reduction (%)") + 
  theme_light()

```


## Final Figures

```{r}
theme_set(theme_light() + 
            theme(plot.title = element_text(size = 24))
)
```


### Box Plots and Line Charts
```{r}
boxplot <- full_dataset_up %>%
  filter(rounded_size <= 1.025) %>%
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  ggplot(aes(rounded_size, slope_reduction * SCALE_FACTOR)) +
  geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0))) +
  scale_x_continuous(breaks = c(0:10)) +
  coord_cartesian(xlim = c(1,10.5), ylim = c(0, 500)) + 
  labs(x = "Battery Size (GWh)", 
       y = "Ramp Rate Reduction (MW/hr)",
       title = "Range of Ramp Rate Reductions",
       subtitle = "Different battery sizes have varying distributions of ramp rate reduction.*",
       caption = "Based on 2012 Texas total wind generation.\n *Note: some outliers were not plotted."
       )

boxplot

ggsave(plot = boxplot, filename = str_c("boxplot", ".png"),
       width = 7, height = 5, dpi = 900, units = "in", path = path_fig, device = "png")



median_plot <- full_medians_up %>% 
  ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median_percent * 100)) + 
  geom_ribbon(aes(ymin = q10_percent * 100, 
                  ymax = q90_percent * 100),
              alpha = 0.3) +
  geom_line(size = 1.5) + 
  coord_cartesian(xlim = c(0,50)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Battery Size (GWh)", 
       y = "Median Ramp Rate Reduction (%)",
       title = "Relative Ramp Reduction",
       subtitle = "A 42 GWh battery will completely flatten all detected wind up-ramp events.",
       caption = "Based on 2012 Texas total wind generation.\nThe grey region signifies ramp reductions between the 10th and 90th percentile for each battery size."
       )

median_plot

ggsave(plot = median_plot, filename = str_c("median_plot", ".png"),
       width = 7, height = 5, dpi = 900, units = "in", path = path_fig, device = "png")
```

### ECDF
```{r}
ecdf_batt1 <- max_events_up %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == 1) %>% 
  .$slope_reduction %>% 
  ecdf(.)

ecdf_batt2 <- max_events_up %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == 5) %>% 
  .$slope_reduction %>% 
  ecdf(.)

ecdf_batt3 <- max_events_up %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == 10) %>% 
  .$slope_reduction %>% 
  ecdf(.)

y_breaks = seq(0, 1, 0.25)
y_labels = c("0%", "25%", "50%", "75%", "100%")

ecdf_plot <- full_dataset_up %>% 
  distinct(slope_reduction) %>% 
  mutate(`1 GWh` = ecdf_batt1(slope_reduction),
         `5 GWh` = ecdf_batt2(slope_reduction),
         `10 GWh` = ecdf_batt3(slope_reduction)
         ) %>% 
  gather(-slope_reduction, key = "Battery Size", value = "ecdf") %>% 
  mutate(`Battery Size` = as.factor(`Battery Size`),
         `Battery Size` = fct_relevel(`Battery Size`,
                                             c("1 GWh", "5 GWh",
                                                 "10 GWh"))
         ) %>%
  ggplot(aes(slope_reduction * SCALE_FACTOR, 1 - ecdf, color = `Battery Size`)) +
  geom_line(size = 1.5) +
  coord_cartesian(xlim = c(0, 500)) +
  scale_y_continuous(breaks = y_breaks, labels = y_labels) +
  labs(x = "Minimum Ramp Rate Reduction (MW/hr)", 
       y = "Percent of Ramps Above Threshold",
       title = "Ramp Rate Reduction Ability",
       subtitle = "Larger batteries can smooth ramps to a greater degree.",
       caption = "Based on 2012 Texas total wind generation."
       ) +
  theme(legend.position = "bottom", legend.text = element_text(size = 12))
  # guides(color = guide_legend(title = element_text(size = 14)))

ecdf_plot

ggsave(plot = ecdf_plot, filename = str_c("ecdf_plot", ".png"),
       width = 6.5, height = 6, dpi = 900, units = "in", path = path_fig, device = "png") 
```

 - make ECDF plots with fixed battery size

