---
title: "CEE 272R Case 2 Exploration"
author: "Sam Kramer"
date: "6/3/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(forcats)
library(stringr)

path_fig = "~/Desktop/"

SCALE_FACTOR = 10000
```

## Functions

This function rounds battery sizes to the nearest 0.005, or 50 MWh. 
```{r}
round_frac <- function(x) {
  x <- x * 200
  x <- ceiling(x)
  x <- x / 200
  x
}
```


## Read data
```{r}
setwd("~/Documents/School/Spring/CEE 272R/Project/CEE272R/case_2/results/")

up_none_raw <- read_csv("battery_power_up_none.csv", col_names = FALSE)
up_5_raw <- read_csv("battery_power_up_five.csv", col_names = FALSE)
up_10_raw <- read_csv("battery_power_up_ten.csv", col_names = FALSE)
up_20_raw <- read_csv("battery_power_up_twenty.csv", col_names = FALSE)


down_none_raw <- read_csv("battery_power_down_none.csv", col_names = FALSE)
down_5_raw <- read_csv("battery_power_down_five.csv", col_names = FALSE)
down_10_raw <- read_csv("battery_power_down_ten.csv", col_names = FALSE)
down_20_raw <- read_csv("battery_power_down_twenty.csv", col_names = FALSE)


ramp_rates_up_raw <- read_csv("Ramp_Rates_Up.csv",
                              col_names = FALSE)
ramp_rates_down_raw <- read_csv("Ramp_Rates_Down.csv",
                                col_names = FALSE)
```

## Tidy the data

#### Up Ramps
```{r}
# Create column names (i.e. ramp reduction amounts) for the data frame
reduction_up <- seq(0,(ncol(up_none_raw) - 1) * 0.0001,
                    length.out = ncol(up_none_raw))

colnames(up_none_raw) <- reduction_up
colnames(up_5_raw) <- reduction_up
colnames(up_10_raw) <- reduction_up
colnames(up_20_raw) <- reduction_up


# Add a column for "event number" and tidy the data
up_none <- up_none_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("u", event_number)) %>% 
  select(-event_number) %>% 
  rename(`0` = battery_size)

up_5 <- up_5_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("u", event_number)) %>% 
  select(-event_number) %>% 
  rename(`5` = battery_size)

up_10 <- up_10_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("u", event_number)) %>% 
  select(-event_number) %>% 
  rename(`10` = battery_size)

up_20 <- up_20_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("u", event_number)) %>% 
  select(-event_number) %>% 
  rename(`20` = battery_size)


up_ramps <- up_none %>% 
  full_join(up_5, by = c("slope_reduction", "id")) %>% 
  full_join(up_10, by = c("slope_reduction", "id")) %>% 
  full_join(up_20, by = c("slope_reduction", "id")) %>% 
  gather(-slope_reduction, -id, key = "E_to_P", value = "battery_size") %>% 
  mutate(E_to_P = as.integer(E_to_P))


up_ramps
```

#### Down Ramps
```{r}
reduction_down <- seq(0,(ncol(down_none_raw) - 1) * 0.0001,
                      length.out = ncol(down_none_raw))
colnames(down_none_raw) <- reduction_down
colnames(down_5_raw) <- reduction_down
colnames(down_10_raw) <- reduction_down
colnames(down_20_raw) <- reduction_down

down_none <- down_none_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("d", event_number)) %>% 
  select(-event_number) %>% 
  rename(`0` = battery_size)

down_5 <- down_5_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("d", event_number)) %>% 
  select(-event_number) %>% 
  rename(`5` = battery_size)

down_10 <- down_10_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("d", event_number)) %>% 
  select(-event_number) %>% 
  rename(`10` = battery_size)

down_20 <- down_20_raw %>% 
  mutate(event_number = row_number()) %>%
  gather(-event_number, key = "slope_reduction", value = "battery_size") %>% 
  mutate(slope_reduction = as.numeric(slope_reduction),
         id = str_c("d", event_number)) %>% 
  select(-event_number) %>% 
  rename(`20` = battery_size)

down_ramps <- down_none %>% 
  full_join(down_5, by = c("slope_reduction", "id")) %>% 
  full_join(down_10, by = c("slope_reduction", "id")) %>% 
  full_join(down_20, by = c("slope_reduction", "id")) %>% 
  gather(-slope_reduction, -id, key = "E_to_P", value = "battery_size") %>% 
  mutate(E_to_P = as.integer(E_to_P))


down_ramps
```



### Clean true ramp rate data

#### Up Ramps
```{r}
ramp_rates_up <- ramp_rates_up_raw

colnames(ramp_rates_up) <- c("ramp_rate")

ramp_rates_up <- ramp_rates_up %>%
  mutate(event_number = row_number(),
         id = str_c("u", event_number)) %>% 
  select(-event_number)

ramp_rates_up
```

#### Down Ramps
```{r}
ramp_rates_down <- ramp_rates_down_raw

colnames(ramp_rates_down) <- c("ramp_rate")

ramp_rates_down <- ramp_rates_down %>%
  mutate(event_number = row_number(),
         id = str_c("d", event_number)) %>% 
  select(-event_number)

ramp_rates_down
```

## Create Table for All Ramps

```{r}
all_ramps <- up_ramps %>% 
  rename(up = battery_size) %>% 
  full_join(down_ramps %>% rename(down = battery_size),
            by = c("id", "slope_reduction","E_to_P")) %>% 
  gather(up, down, key = "direction", value = "battery_size")

all_ramps
```


```{r}
ramp_rates_all <- ramp_rates_up %>%
  rename(up = ramp_rate) %>%
  full_join(ramp_rates_down %>% rename(down = ramp_rate),
            by = "id") %>%
  gather(-id, key = "direction", value = "ramp_rate") %>% 
  filter(!is.na(ramp_rate))

ramp_rates_all
```


## Add variables

### Calculate percentage reduction

#### Up Ramps
```{r}
up_ramps <- up_ramps %>%
  left_join(ramp_rates_up, by = "id") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate) %>%
  select(id, slope_reduction, percent_reduction, battery_size, everything())

up_ramps
```

#### Down Ramps
```{r}
down_ramps <- down_ramps %>% 
  left_join(ramp_rates_down, by = "id") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate) %>%
  select(id, slope_reduction, percent_reduction, battery_size, everything())

down_ramps
```


#### All Ramps
```{r}
all_ramps <- all_ramps %>% 
  left_join(ramp_rates_all, 
            by = c("id", "direction")) %>% 
  filter(!is.na(ramp_rate)) %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate) %>% 
  select(id, direction, battery_size, slope_reduction, percent_reduction, 
         everything())

all_ramps
```


### Round battery sizes

#### Up Ramps
```{r}
# Round the battery sizes
up_ramps <- up_ramps %>% 
  mutate(rounded_size = round_frac(battery_size))

up_ramps
```

#### Down Ramps
```{r}
# Round the battery sizes
down_ramps <- down_ramps %>% 
  mutate(rounded_size = round_frac(battery_size))

down_ramps
```

#### All Ramps
```{r}
all_ramps <- all_ramps %>% 
  mutate(rounded_size = round_frac(battery_size))

all_ramps
```


### Find maximum slope reduction for each combination of event and battery size
```{r}
max_events_up <- up_ramps %>% 
  group_by(rounded_size,id, E_to_P) %>% 
  summarize(slope_reduction = max(slope_reduction),
            percent_reduction = max(percent_reduction)
            ) %>% 
  ungroup()

max_events_up
```

```{r}
max_events_down <- down_ramps %>% 
  group_by(rounded_size,id, E_to_P) %>% 
  summarize(slope_reduction = max(slope_reduction),
            percent_reduction = max(percent_reduction)
            ) %>% 
  ungroup()

max_events_down
```


```{r}
max_events_all <- all_ramps %>% 
  group_by(rounded_size, id, direction, E_to_P) %>% 
  summarize(slope_reduction = max(slope_reduction),
            percent_reduction = max(percent_reduction)
            ) %>% 
  ungroup()

max_events_all
```

## Make full dataset for all battery sizes and events

#### Up Ramps
```{r}
full_dataset_up <- max_events_up %>% 
  data_grid(rounded_size, id, E_to_P) %>% 
  left_join(max_events_up, by = c("id", "rounded_size", "E_to_P")) %>% 
  select(id, rounded_size, E_to_P, slope_reduction) %>% 
  arrange(E_to_P, id, rounded_size) %>% 
  mutate(slope_reduction = ifelse(rounded_size == 0, 0, slope_reduction)) %>% 
  fill(slope_reduction) %>% 
  left_join(ramp_rates_up, by = "id") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate)

full_dataset_up


full_medians_up <- full_dataset_up %>% 
  group_by(rounded_size, E_to_P) %>% 
  summarize(median = median(slope_reduction),
            median_percent = median(percent_reduction),
            q10 = quantile(slope_reduction, 0.1),
            q90 = quantile(slope_reduction, 0.9),
            q10_percent = quantile(percent_reduction, 0.1),
            q90_percent = quantile(percent_reduction, 0.9)
  ) %>% 
  ungroup()

full_medians_up
```

```{r}
full_dataset_down <- max_events_down %>% 
  data_grid(rounded_size, id, E_to_P) %>% 
  full_join(max_events_down, by = c("id", "rounded_size", "E_to_P")) %>% 
  select(id, rounded_size, E_to_P, slope_reduction) %>% 
  arrange(E_to_P, id, rounded_size) %>% 
  mutate(slope_reduction = ifelse(rounded_size == 0, 0, slope_reduction)) %>% 
  fill(slope_reduction) %>% 
  left_join(ramp_rates_down, by = "id") %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate)

full_dataset_down


full_medians_down <- full_dataset_down %>% 
  group_by(rounded_size, E_to_P) %>% 
  summarize(median = median(slope_reduction),
            median_percent = median(percent_reduction),
            q10 = quantile(slope_reduction, 0.1),
            q90 = quantile(slope_reduction, 0.9),
            q10_percent = quantile(percent_reduction, 0.1),
            q90_percent = quantile(percent_reduction, 0.9)
  ) %>% 
  ungroup()

full_medians_down
```


```{r}
full_dataset <- full_dataset_up %>% 
  mutate(direction = "up") %>% 
  full_join(full_dataset_down %>% 
              mutate(direction = "down"), 
            by = c("id", "rounded_size", "slope_reduction",
                   "ramp_rate", "percent_reduction", "direction", "E_to_P")) %>% 
  select(id, direction, rounded_size, E_to_P, slope_reduction, percent_reduction, 
         ramp_rate, everything())

nrow(full_dataset_down)
nrow(full_dataset_up)
nrow(full_dataset_down) + nrow(full_dataset_up)
nrow(full_dataset)


full_medians <- full_dataset %>%
  group_by(direction, rounded_size, E_to_P) %>% 
  summarize(median = median(slope_reduction),
            median_percent = median(percent_reduction),
            q10 = quantile(slope_reduction, 0.1),
            q90 = quantile(slope_reduction, 0.9),
            q10_percent = quantile(percent_reduction, 0.1),
            q90_percent = quantile(percent_reduction, 0.9)) %>% 
  ungroup()

full_dataset
full_medians
```


### Calculate statistics for all ramps together
```{r}
all_ramps_full <- max_events_all %>% 
  data_grid(rounded_size, id, E_to_P) %>% 
  full_join(max_events_all %>% select(-direction), by = c("id", "rounded_size", "E_to_P")) %>% 
  select(id, rounded_size, E_to_P, slope_reduction) %>% 
  arrange(E_to_P, id, rounded_size) %>%
  mutate(slope_reduction = ifelse(rounded_size == 0, 0, slope_reduction)) %>% 
  fill(slope_reduction) %>% 
  left_join(ramp_rates_all %>% select(-direction), by = "id") %>% 
  filter(!is.na(ramp_rate)) %>% 
  mutate(percent_reduction = slope_reduction / ramp_rate)

# max_events_down %>% 
#   data_grid(rounded_size, id, E_to_P) %>% 
#   full_join(max_events_down, by = c("id", "rounded_size", "E_to_P")) %>% 
#   select(id, rounded_size, E_to_P, slope_reduction) %>% 
#   arrange(E_to_P, id, rounded_size) %>% 
#   mutate(slope_reduction = ifelse(rounded_size == 0, 0, slope_reduction)) %>% 
#   fill(slope_reduction) %>% 
#   left_join(ramp_rates_down, by = "id") %>% 
#   mutate(percent_reduction = slope_reduction / ramp_rate)

all_ramps_full_medians <- all_ramps_full %>% 
  group_by(E_to_P, rounded_size) %>% 
  summarize(median = median(slope_reduction, na.rm = T),
            median_percent = median(percent_reduction, na.rm = T),
            q10 = quantile(slope_reduction, 0.1, na.rm = T),
            q90 = quantile(slope_reduction, 0.9, na.rm = T),
            q10_percent = quantile(percent_reduction, 0.1, na.rm = T),
            q90_percent = quantile(percent_reduction, 0.9, na.rm = T)
  ) %>% 
  ungroup()
```

```{r}
missing <- all_ramps_full %>% 
  anti_join(full_dataset, by = c("id", "rounded_size", "E_to_P"))
```

```{r}
missing

missing %>% 
  group_by(id) %>% 
  summarize(rate = max(ramp_rate)) %>% 
  ggplot(aes(rate)) + 
  geom_histogram(binwidth = 0.001)

# full_dataset %>% 
#   filter(id == "u99")

ramp_rates_all %>% 
  ggplot(aes(ramp_rate)) + 
  geom_histogram(binwidth = 0.001)

missing %>% 
  group_by(id) %>% 
  summarize(rate = max(ramp_rate)) %>% 
  anti_join(ramp_rates_all, by = c("id"))

full_dataset_up %>% 
  filter(id == "u99") %>% 
  arrange(desc(rounded_size))
```







```{r}
full_medians_down %>% 
  filter(E_to_P == 0) %>% 
  mutate(E_to_P = as.factor(E_to_P)) %>% 
  ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median_percent * 100)) + 
  geom_ribbon(aes(ymin = q10_percent * 100, 
                  ymax = q90_percent * 100),
              alpha = 0.3) +
  # facet_grid(E_to_P~direction) +
  # facet_wrap(~direction) +
  # geom_line(size = 1.5) + 
  geom_line() + 
  coord_cartesian(xlim = c(0,75)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 75, 15)) +
  labs(x = "Battery Size (GWh)", 
       y = "Median Ramp Rate Reduction (%)",
       title = "Relative Ramp Reduction",
       subtitle = "A 42 GWh battery will completely flatten all detected wind up-ramp events.",
       caption = "Based on 2012 Texas total wind generation.\nThe grey region signifies ramp reductions between the 10th and 90th percentile for each battery size."
       )
```








## Final Figures

```{r}
theme_set(theme_light() +
            theme(#panel.grid.major.x = element_blank(),
                  panel.grid.minor.x = element_blank(),
                  panel.grid.minor.y = element_blank(),
                  axis.title = element_text(size = 13),
                  axis.text = element_text(size = 11),
                  strip.text.x = element_text(size = 13)) +
            theme(legend.position = 'bottom',
                  legend.text = element_text(size = 12),
                  legend.title = element_text(size = 13)) +
            theme(plot.title = element_text(size = 26))
          )
```



### Box Plots

```{r}
boxplot_final <- all_ramps_full %>%
  filter(rounded_size <= 1.025) %>%
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000,
         E_to_P = as.factor(E_to_P)) %>% 
  ggplot(aes(rounded_size, slope_reduction * SCALE_FACTOR)) +
  geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0)),
               outlier.alpha = 0.01) +
  # facet_grid(E_to_P~direction) + 
  scale_x_continuous(breaks = c(0:10)) +
  coord_cartesian(xlim = c(1,10.5), ylim = c(0, 300)) +
  labs(x = "Battery Size (GWh)", 
       y = "Ramp Rate Reduction (MW/hr)")

boxplot_final

ggsave("boxplot_final.png", plot = boxplot_final, device = "png", 
      path = path_fig, width = 6, height = 4, units = "in", dpi = 900)
```


### Line Chart

```{r}
median_final <- all_ramps_full_medians %>% 
  mutate(E_to_P = as.factor(E_to_P),
         E_to_P = fct_recode(E_to_P, `0 hr` = "0", `5 hr` = "5", 
                             `10 hr` = "10", `20 hr` = "20")) %>% 
  ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median_percent * 100)) + 
  # geom_line(aes(color = E_to_P)) +
  geom_line(aes(color = E_to_P), size = 1.2) +
  # facet_wrap(~direction) + 
  coord_cartesian(xlim = c(0,50)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 50, 10)) +
  labs(x = "Battery Size (GWh)", 
       y = "Median Ramp Rate Reduction (%)",
       color = "Energy to Power Ratio"
       )

median_final

ggsave("median_final.png", plot = median_final, device = "png", 
      path = path_fig, width = 6, height = 4, units = "in", dpi = 900)
```



```{r}
# median_plot <- full_medians %>% 
#   mutate(E_to_P = as.factor(E_to_P),
#          direction = ifelse(direction != "down", "Up Ramps", "Down Ramps")
#          ) %>% 
#   ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median_percent * 100)) + 
#   geom_line(aes(color = E_to_P), size = 1.2) +
#   facet_wrap(~direction) + 
#   coord_cartesian(xlim = c(0,75)) +
#   scale_y_continuous(breaks = seq(0, 100, 10)) +
#   scale_x_continuous(breaks = seq(0, 75, 15)) +
#   labs(x = "Battery Size (GWh)", 
#        y = "Median Ramp Rate Reduction (%)",
#        title = "Relative Ramp Reduction",
#        color = "Energy to Power Ratio\n(GWh/GW)"
#        # subtitle = "A 42 GWh battery will completely flatten all detected wind up-ramp events.",
#        # caption = "Based on 2012 Texas total wind generation.\nThe grey region signifies ramp reductions between the 10th and 90th percentile for each battery size."
#        ) + 
#     theme(legend.position = "bottom", legend.text = element_text(size = 12))
# 
# 
# median_plot
# 
# # ggsave(plot = median_plot, filename = str_c("median_plot", ".png"),
# #        width = 7, height = 5, dpi = 900, units = "in", path = path_fig, device = "png")
```

### ECDF

#### Baseline case: no restriction on charge/discharge rate
```{r}
E2P_FIXED = 5

ecdf_all_1 <- max_events_all %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == 1,
         E_to_P == E2P_FIXED) %>% 
  .$slope_reduction %>% 
  ecdf(.)

ecdf_all_5 <- max_events_all %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == 5,
         E_to_P == E2P_FIXED) %>% 
  .$slope_reduction %>% 
  ecdf(.)

ecdf_all_10 <- max_events_all %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == 10,
         E_to_P == E2P_FIXED) %>% 
  .$slope_reduction %>% 
  ecdf(.)
```

```{r}
y_breaks = seq(0, 1, 0.25)
y_labels = c("0%", "25%", "50%", "75%", "100%")

ecdf_batt_final <- all_ramps_full %>%
  filter(rounded_size == E2P_FIXED) %>% 
  distinct(slope_reduction) %>%
  mutate(`1 GWh` = ecdf_all_1(slope_reduction),
         `5 GWh` = ecdf_all_5(slope_reduction),
         `10 GWh` = ecdf_all_10(slope_reduction)
         ) %>%
  gather(-slope_reduction, key = "Battery Size", value = "ecdf") %>%
  mutate(`Battery Size` = as.factor(`Battery Size`),
         `Battery Size` = fct_relevel(`Battery Size`, c("1 GWh", "5 GWh", "10 GWh"))
         ) %>%
  ggplot(aes(slope_reduction * SCALE_FACTOR, 1 - ecdf)) +
  geom_line(aes(color = `Battery Size`),  size = 1.5) +
  coord_cartesian(xlim = c(0, 400)) +
  scale_y_continuous(breaks = y_breaks, labels = y_labels) +
  labs(x = "Minimum Ramp Rate Reduction (MW/hr)",
       y = "Percent of Ramps Above Threshold"
       ) +
  theme(legend.position = "bottom", legend.text = element_text(size = 12))

ecdf_batt_final

ggsave("ecdf_batt_final.png", plot = ecdf_batt_final, device = "png", 
      path = path_fig, width = 6, height = 4, units = "in", dpi = 900)
```



#### Fixed battery size, variable E to P ratio

```{r}
BATT_SIZE = 5

ecdf_all_e2p0 <- max_events_all %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == BATT_SIZE,
         E_to_P == 0) %>% 
  .$slope_reduction %>% 
  ecdf(.)

ecdf_all_e2p5 <- max_events_all %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == BATT_SIZE,
         E_to_P == 5) %>% 
  .$slope_reduction %>% 
  ecdf(.)

ecdf_all_e2p10 <- max_events_all %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == BATT_SIZE,
         E_to_P == 10) %>% 
  .$slope_reduction %>% 
  ecdf(.)

ecdf_all_e2p20 <- max_events_all %>% 
  mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
  filter(rounded_size == BATT_SIZE,
         E_to_P == 20) %>% 
  .$slope_reduction %>% 
  ecdf(.)
```

```{r}
y_breaks = seq(0, 1, 0.25)
y_labels = c("0%", "25%", "50%", "75%", "100%")

ecdf_e2p_final <- all_ramps_full %>%
  filter(rounded_size == BATT_SIZE) %>% 
  distinct(slope_reduction) %>%
  mutate(`0` = ecdf_all_e2p0(slope_reduction),
         `5` = ecdf_all_e2p5(slope_reduction),
         `10` = ecdf_all_e2p10(slope_reduction),
         `20` = ecdf_all_e2p20(slope_reduction)
         ) %>%
  gather(-slope_reduction, key = "Energy to Power Ratio", value = "ecdf") %>%
  mutate(`Energy to Power Ratio` = 
           as.integer(`Energy to Power Ratio`),
          `Energy to Power Ratio` = 
           as.factor(`Energy to Power Ratio`),
         `Energy to Power Ratio` = 
           fct_recode(`Energy to Power Ratio`,
                      `0 hr` = "0", `5 hr` = "5", `10 hr` = "10", `20 hr` = "20")

         ) %>%
  ggplot(aes(slope_reduction * SCALE_FACTOR, 1 - ecdf)) +
  geom_line(aes(color = `Energy to Power Ratio`),  size = 1.5) +
  coord_cartesian(xlim = c(0, 400)) +
  scale_y_continuous(breaks = y_breaks, labels = y_labels) +
  labs(x = "Minimum Ramp Rate Reduction (MW/hr)",
       y = "Percent of Ramps Above Threshold"
       ) +
  theme(legend.position = "bottom", legend.text = element_text(size = 12))

ecdf_e2p_final

ggsave("ecdf_e2p_final.png", plot = ecdf_e2p_final, device = "png", 
      path = path_fig, width = 6, height = 4, units = "in", dpi = 900)
```


```{r}
# y_breaks = seq(0, 1, 0.25)
# y_labels = c("0%", "25%", "50%", "75%", "100%")
# 
# full_dataset %>%
#   filter(E_to_P == 0) %>% 
#   distinct(direction, slope_reduction) %>%
#   mutate(`1 GWh` = ifelse(direction == "up", ecdf_up_1(slope_reduction),
#                       ecdf_down_1(slope_reduction)),
#          `5 GWh` = ifelse(direction == "up", ecdf_up_5(slope_reduction),
#                       ecdf_down_5(slope_reduction)),
#          `10 GWh` = ifelse(direction == "up", ecdf_up_10(slope_reduction),
#                       ecdf_down_10(slope_reduction))) %>%
#   gather(-slope_reduction,-direction, key = "Battery Size", value = "ecdf") %>%
#   mutate(`Battery Size` = as.factor(`Battery Size`),
#          `Battery Size` = fct_relevel(`Battery Size`,
#                                              c("1 GWh", "5 GWh",
#                                                  "10 GWh")),
#           direction = ifelse(direction != "down", "Up Ramps", "Down Ramps")
# 
#          ) %>%
#   ggplot(aes(slope_reduction * SCALE_FACTOR, 1 - ecdf, color = `Battery Size`)) +
#   geom_line(size = 1.5) +
#   # facet_wrap(~direction) +
#   coord_cartesian(xlim = c(0, 500)) +
#   scale_y_continuous(breaks = y_breaks, labels = y_labels) +
#   labs(x = "Minimum Ramp Rate Reduction (MW/hr)",
#        y = "Percent of Ramps Above Threshold",
#        title = "Ramp Rate Reduction Ability",
#        subtitle = "Larger batteries can smooth ramps to a greater degree.",
#        caption = "Based on 2012 Texas total wind generation."
#        ) +
#   theme(legend.position = "bottom", legend.text = element_text(size = 12))
```




```{r}
# BATT_SIZE = 5
# 
# ecdf_up_e2p0 <- max_events_up %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_up_e2p5 <- max_events_up %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 5) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_up_e2p10 <- max_events_up %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 10) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_up_e2p20 <- max_events_up %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 20) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_down_e2p0 <- max_events_down %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_down_e2p5 <- max_events_down %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 5) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_down_e2p10 <- max_events_down %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 10) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_down_e2p20 <- max_events_down %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == BATT_SIZE,
#          E_to_P == 20) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
```


```{r}
# y_breaks = seq(0, 1, 0.25)
# y_labels = c("0%", "25%", "50%", "75%", "100%")
# 
# full_dataset %>%
#   filter(rounded_size == BATT_SIZE) %>% 
#   distinct(direction, slope_reduction) %>%
#   mutate(`0 GWh/GW` = ifelse(direction == "up", ecdf_up_e2p0(slope_reduction),
#                       ecdf_down_e2p0(slope_reduction)),
#          `5 GWh/GW` = ifelse(direction == "up", ecdf_up_e2p5(slope_reduction),
#                       ecdf_down_e2p5(slope_reduction)),
#          `10 GWh/GW` = ifelse(direction == "up", ecdf_up_e2p10(slope_reduction),
#                       ecdf_down_e2p10(slope_reduction)),
#          `20 GWh/GW` = ifelse(direction == "up", ecdf_up_e2p20(slope_reduction),
#                       ecdf_down_e2p20(slope_reduction))
#          ) %>%
#   gather(-slope_reduction,-direction, key = "Power to Energy Ratio", value = "ecdf") %>%
#   mutate(`Power to Energy Ratio` = as.factor(`Power to Energy Ratio`),
#          `Power to Energy Ratio` = fct_relevel(`Power to Energy Ratio`,
#                                              c("0 GWh/GW", "5 GWh/GW",
#                                                  "10 GWh/GW", "20 GWh/GW")),
#           direction = ifelse(direction != "down", "Up Ramps", "Down Ramps")
# 
#          ) %>%
#   ggplot(aes(slope_reduction * SCALE_FACTOR, 1 - ecdf, color = `Power to Energy Ratio`)) +
#   geom_line(size = 1.5) +
#   facet_wrap(~direction) +
#   coord_cartesian(xlim = c(0, 400)) +
#   scale_y_continuous(breaks = y_breaks, labels = y_labels) +
#   labs(x = "Minimum Ramp Rate Reduction (MW/hr)",
#        y = "Percent of Ramps Above Threshold",
#        title = "Effect of Charge/Discharge Rate",
#        subtitle = "Lower energy-to-power ratios allow for inreased smoothing with a fixed battery size.",
#        caption = "Based on 2012 Texas total wind generation. Battery size is fixed at 5 GWh"
#        ) +
#   theme(legend.position = "bottom", legend.text = element_text(size = 12))
```




```{r}
# ecdf_up_1 <- max_events_up %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == 1,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_up_5 <- max_events_up %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == 5,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_up_10 <- max_events_up %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == 10,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_down_1 <- max_events_down %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == 1,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_down_5 <- max_events_down %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == 5,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)
# 
# ecdf_down_10 <- max_events_down %>% 
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   filter(rounded_size == 10,
#          E_to_P == 0) %>% 
#   .$slope_reduction %>% 
#   ecdf(.)

```








```{r}
# all_ramps_full_medians %>% 
#   # filter(E_to_P != 5) %>% 
#   # mutate(E_to_P = as.factor(E_to_P),
#   #        direction = ifelse(direction != "down", "Up Ramps", "Down Ramps")
#   #        ) %>% 
#   mutate(E_to_P = as.factor(E_to_P)) %>% 
#   ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median * SCALE_FACTOR)) + 
#   # geom_ribbon(aes(ymin = q10_percent * 100, 
#   #                 ymax = q90_percent * 100),
#   #             alpha = 0.3) +
#   # facet_grid(E_to_P~direction) +
#   geom_line(aes(color = E_to_P), size = 1.2) +
#   # facet_wrap(~direction) + 
#   # geom_line() +
#   coord_cartesian(xlim = c(0,75)) +
#   # scale_y_continuous(breaks = seq(0, 100, 10)) +
#   scale_x_continuous(breaks = seq(0, 75, 15)) +
#   labs(x = "Battery Size (GWh)", 
#        y = "Median Ramp Rate Reduction (MW/hr)",
#        # title = "Relative Ramp Reduction",
#        color = "Energy to Power Ratio\n(GWh/GW)"
#        # subtitle = "A 42 GWh battery will completely flatten all detected wind up-ramp events.",
#        # caption = "Based on 2012 Texas total wind generation.\nThe grey region signifies ramp reductions between the 10th and 90th percentile for each battery size."
#        ) + 
#     theme(legend.position = "bottom", legend.text = element_text(size = 12))
```

## Data Exploration


## Plot data
```{r}
# up_ramps %>% 
#   ggplot(aes(battery_size * SCALE_FACTOR)) + 
#   geom_histogram(binwidth = 0.05 * SCALE_FACTOR)
# 
# all_ramps %>% 
#   ggplot(aes(battery_size * SCALE_FACTOR)) + 
#   geom_histogram(binwidth = 0.05 * SCALE_FACTOR) + 
#   facet_wrap(~direction)

# up_ramps %>% 
#   filter(slope_reduction == 1*10^-4,
#          battery_size < quantile(battery_size, 0.9, na.rm = TRUE)) %>% 
#   ggplot(aes(battery_size)) + 
#   geom_histogram(bins = 100)

# up_ramps %>% 
#   filter(slope_reduction %in% quantile(up_ramps$slope_reduction)) %>% 
#   ggplot(aes(battery_size)) + 
#   geom_freqpoly(aes(color = factor(slope_reduction)), binwidth = 0.05)
```


```{r}
# up_ramps %>% 
#   filter(!is.nan(battery_size)) %>% 
#   ggplot(aes(slope_reduction * SCALE_FACTOR)) + 
#   geom_histogram(binwidth = 0.0001*SCALE_FACTOR) + 
#   labs(x = "Ramp Rate Reduction (MW/hr)", y = "Count")
# 
# up_ramps %>% 
#   filter(!is.nan(battery_size)) %>% 
#   ggplot(aes(slope_reduction * SCALE_FACTOR, ..density..)) + 
#   geom_histogram(binwidth = 0.0001*SCALE_FACTOR) + 
#   labs(x = "Ramp Rate Reduction (MW/hr)", y = "Fraction of Events")

# all_ramps %>% 
#   filter(!is.nan(battery_size)) %>% 
#   ggplot(aes(slope_reduction * SCALE_FACTOR)) + 
#   geom_histogram(binwidth = 0.0001*SCALE_FACTOR) + 
#   labs(x = "Ramp Rate Reduction (MW/hr)", y = "Count") + 
#   facet_wrap(~direction)
# 
# all_ramps %>% 
#   filter(!is.nan(battery_size)) %>% 
#   ggplot(aes(slope_reduction * SCALE_FACTOR, ..density..)) + 
#   geom_histogram(binwidth = 0.0001*SCALE_FACTOR) + 
#   labs(x = "Ramp Rate Reduction (MW/hr)", y = "Fraction of Events") + 
#   facet_wrap(~direction)

# up_ramps %>% 
#   filter(!is.nan(battery_size))
```


```{r}
# full_dataset_up %>%
#   filter(rounded_size <= 1.025) %>%
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   ggplot(aes(rounded_size, slope_reduction * SCALE_FACTOR)) +
#   # geom_boxplot(aes(group = cut_number(rounded_size, 10)), outlier.alpha = 0) +
#   geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0)), outlier.alpha = 0) +
#   scale_x_continuous(breaks = c(0:10)) +
#   # coord_cartesian(xlim = c(1,10.5), ylim = c(0, 450)) + 
#   labs(x = "Battery Size (GWh)", y = "Ramp Rate Reduction (MW/hr)") +
#   theme_light()
# 
# full_dataset_up %>%
#   filter(rounded_size <= 1.025) %>%
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   ggplot(aes(rounded_size, percent_reduction * 100)) +
#   # geom_boxplot(aes(group = cut_number(rounded_size, 10))) +
#   geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0)), outlier.alpha = 0) +
#   # coord_cartesian(xlim = c(1,10.5)) +
#   scale_x_continuous(breaks = c(0:10)) +
#   labs(x = "Battery Size (GWh)", y = "Ramp Rate Reduction (%)") +
#   theme_light()
# 
# full_medians_up %>% 
#   ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median * SCALE_FACTOR)) + 
#   geom_ribbon(aes(ymin = q10 * SCALE_FACTOR, ymax = q90 * SCALE_FACTOR),
#               alpha = 0.3) +
#   geom_line() + 
#   # coord_cartesian(xlim = c(0, 50)) +
#   labs(x = "Battery Size (GWh)", y = "Median Ramp Rate Reduction (MW/hr)") +
#   theme_light()
# 
# full_medians_up %>% 
#   ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median_percent * 100)) + 
#   geom_ribbon(aes(ymin = q10_percent * 100, 
#                   ymax = q90_percent * 100),
#               alpha = 0.3) +
#   geom_line() + 
#   # coord_cartesian(xlim = c(0,50)) +
#   scale_y_continuous(breaks = seq(0, 100, 10)) +
#   labs(x = "Battery Size (GWh)", y = "Median Ramp Rate Reduction (%)") + 
#   theme_light()

```


```{r}
# full_dataset_down %>%
#   filter(rounded_size <= 1.025) %>%
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   ggplot(aes(rounded_size, slope_reduction * SCALE_FACTOR)) +
#   # geom_boxplot(aes(group = cut_number(rounded_size, 10)), outlier.alpha = 0) +
#   geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0)), outlier.alpha = 0) +
#   scale_x_continuous(breaks = c(0:10)) +
#   coord_cartesian(xlim = c(1,10.5), ylim = c(0, 450)) + 
#   labs(x = "Battery Size (GWh)", y = "Ramp Rate Reduction (MW/hr)") +
#   theme_light()
# 
# full_dataset_down %>%
#   filter(rounded_size <= 1.025) %>%
#   mutate(rounded_size = rounded_size * SCALE_FACTOR / 1000) %>% 
#   ggplot(aes(rounded_size, percent_reduction * 100)) +
#   # geom_boxplot(aes(group = cut_number(rounded_size, 10))) +
#   geom_boxplot(aes(group = cut_width(rounded_size, 1, center = 0)), outlier.alpha = 0) +
#   coord_cartesian(xlim = c(1,10.5)) +
#   scale_x_continuous(breaks = c(0:10)) +
#   labs(x = "Battery Size (GWh)", y = "Ramp Rate Reduction (%)") +
#   theme_light()
# 
# full_medians_down %>% 
#   ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median * SCALE_FACTOR)) + 
#   geom_ribbon(aes(ymin = q10 * SCALE_FACTOR, ymax = q90 * SCALE_FACTOR),
#               alpha = 0.3) +
#   geom_line() + 
#   coord_cartesian(xlim = c(0, 50)) +
#   labs(x = "Battery Size (GWh)", y = "Median Ramp Rate Reduction (MW/hr)") +
#   theme_light()
# 
# full_medians_down %>% 
#   ggplot(aes(rounded_size * SCALE_FACTOR / 1000, median_percent * 100)) + 
#   geom_ribbon(aes(ymin = q10_percent * 100, 
#                   ymax = q90_percent * 100),
#               alpha = 0.3) +
#   geom_line() + 
#   coord_cartesian(xlim = c(0,50)) +
#   scale_y_continuous(breaks = seq(0, 100, 10)) +
#   labs(x = "Battery Size (GWh)", y = "Median Ramp Rate Reduction (%)") + 
#   theme_light()

```


